---
output:
  word_document: default
  html_document: default
---
### Clustering  

## Libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
```

```{r message=FALSE, warning=FALSE}
trucks <- read_csv("trucks.csv")

#str(trucks)
```
## Task 1  

There does appear to be some natural clustering in the data. There is a cluster of points traveling less than ~80 miles and speeding for less than ~12% of the time. Another cluster of those that were speeding for more than 12% of the time while driving 80 miles or less. There are 2 more clusters for those traveling 125+ miles. There is a group that was speeding between 0-25% of the trip and another, more scattered cluster for those that were speeding for more than 25% of the trip.   
```{r}
ggplot(trucks, aes(x=Distance, y=Speeding)) + geom_point() 
```

## Task 2  
```{r}
kmeans_recipe = recipe(~ Distance + Speeding, trucks)

trucks_dummy = kmeans_recipe %>%
  step_scale(all_numeric()) %>%
  step_center(all_numeric())

trucks_dummy = prep(trucks_dummy, trucks)

trucks_cleaned = bake(trucks_dummy, trucks)
```

### Task 3  

The clusters are predictable based on the initial scatter plot. With only 2 groups, they are grouped by distance traveled into those above 125 and those under 125 miles.  

```{r}
set.seed(64)
clusters = kmeans(trucks_cleaned, 2)
```

```{r}
trucks = augment(clusters, trucks)
str(trucks)
```
```{r}
ggplot(trucks, aes(x=Distance, y=Speeding, color=factor(.cluster))) + geom_point()
```
  


## Task 4  

4 clusters seem appropriate for this data set based on the visualizations below.  

```{r}
set.seed(412)
clusts =
  tibble(k= 1:8) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusts
```

```{r}
clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>%
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))
```

```{r}
p1 = 
  ggplot(assignments, aes(x=Distance, y=Speeding)) + geom_point(aes(color=.cluster), alpha = 0.8) + facet_wrap(~k)

p1
```

  

## Task 5  

Based on the plot, k = 4 is at the "elbow" of the line and seems to be most appropriate. 

```{r}
ggplot(clusterings, aes(k, tot.withinss)) + geom_line() + geom_point()
```


## Task 6  

These clusters seem most appropriate for this data based on the analysis above. It was predicted based on the initial visualization that 4 clusters could be made from this data. The analysis agrees with the inital prediction.  

```{r}
set.seed(64)
clusters = kmeans(trucks_cleaned, 4)
```

```{r}
trucks = augment(clusters, trucks)
str(trucks)
```

```{r}
ggplot(trucks, aes(x=Distance, y=Speeding, color=factor(.cluster))) + geom_point()
```



