---
output:
  word_document: default
  html_document: default
---
## Model Validation

### Libraries
```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(lubridate)
library(tidymodels)
```

```{r message = FALSE, warning = FALSE}
bike <- read_csv("bike_cleaned.csv")
bike = bike %>% 
  mutate(dteday = mdy(dteday)) %>% 
  mutate_if(is.character, as_factor) %>%
  mutate(hr = as_factor(hr))
#str(bike)
```
### Task 1: Training and Testing Sets  
```{r}
set.seed(1234)
bike_split = initial_split(bike, prob=0.7, strata = count)
train = training(bike_split)
test = testing(bike_split)
```

### Task 2: Examine the data  
**Train -** 13,032 rows  
**Test -** 4,347 rows  
```{r}
#str(train)
#str(test)
```

### Task 3: Linear Regression
**The adjusted R squared value is 0.6223. There were many values that were not significant in this model. Perhaps a lasso model would yield better results but overall, the R squared value is satisfactory  ** 
```{r}
bike_recipe = recipe(count ~ season + mnth + hr + holiday + weekday + temp + weathersit, train) 

lm_model =
  linear_reg() %>%
  set_engine("lm")

lm_wflow =
  workflow() %>%
  add_model(lm_model) %>%
  add_recipe(bike_recipe)

lm_fit = fit(lm_wflow, train)

summary(lm_fit$fit$fit$fit)
```
### Task 4: Testing
**The histogram of the prediction shows a relatively normal distribution**
```{r}
predict_train <- predict(lm_fit, new_data = test)

ggplot(predict_train, aes(x = .pred)) + geom_histogram()
```

### Task 5: Performance
**The R squared value of the test set is 0.6248 which is similar to to that of the previous model (0.6223). The model performed well as the R squared value did not change significantly with the addition of the test data  **
```{r}
lm_fit %>% 
  predict(test) %>%
  bind_cols(test) %>%
  metrics(truth = count, estimate = .pred)
```

